### 2023-11-08 generalExploration.Rmd
moving model function in there own file, adding a package like structure

source("R/population.R")
N=100
Th=400
z=4 #number of neutral traits
ki <- 5
km <- 2
K  <-  ki+km
m=.2 #proba marriage
b=.1 #HG birth rate
r=.2 #farmer scale param
rho=.5 #marriage rule
d=0.01

a=initAdaptiveTraits(ki=ki,km=km)
neutraltraitsParam=initNeutralTraits(z = z)
neutraltraitsParam$s=c(0,0,1,1)

reproduction <- function(p1,p2,tp,i=NULL){
    stopifnot(p1$community==p2$community)
    traits=ifelse(sample(c(0,1),length(tp$s),replace=T),p1$traits,p2$traits)
    if(!is.null(i))id=as.numeric(i)+1
    list(
         id=as.character(id),
         traits=traits,
         age=0,
         community=p1$community,
         sex=sample(c("F","M"),1),
         repro=FALSE,
         partner=-1
    )
}

@@

pos=random2Dgrid(K=K,Gx=100)
initcomus=initialiseCommunities(traits=a,coordinates=pos)
size=table(sapply(population,"[[","community"))
initcomus$size=size
plot(initcomus$coordinates,pch=21,bg=apply(initcomus$adaptivetraits,1,mean)+1,cex=log(initcomus$size))
logging="time"
tstep=1000
popsize=c()
popsum=list()
for(time in 1:tstep){
    if(logging=="time")print(paste("------",time,"------"))
    for(i in sample(1:length(population))){
        ind=tryCatch(population[[i]]$id,error=function(e)NULL)
        if(!is.null(ind)){
            ##ageing

            population[[ind]]$age=population[[ind]]$age+1

            ##repro
            partner=population[[ind]]$partner
            canrepro=population[[ind]]$repro
            if(partner>0 & !canrepro){
                if(logging=="verbose")print(paste("childtest",ind,partner,population[[ind]]$community,population[[partner]]$community))
                population[[partner]]$repro=TRUE #make sure proba of reproduction is unique for all pair 
                ad_tr=initcomus$adaptivetraits[population[[ind]]$community,]
                lambda=b+r*sum(ad_tr)
                nchild=rpois(1,lambda)
                if(nchild>0){
                    for(i in 1:nchild){
                        newborn=reproduction(population[[ind]],population[[partner]],neutraltraitsParam,population[[length(population)]]$id)
                        if(logging=="demo")print(paste("newborn in ",newborn$community))
                        population[[newborn$id]]=newborn
                        initcomus$size[newborn$community]=initcomus$size[newborn$community]+1
                    }
                }
            }
            population[[ind]]$repro=FALSE

            ##marriage
            if(population[[ind]]$age>18 & population[[ind]]$partner <0){
                #print("marriage attempt")
                if(runif(1)<m){
                    potential=sapply(population,function(i,s)if(i$age>18 & i$partner<0 & i$sex != s )return(i$id),s=population[[ind]]$sex)
                    potential=unlist(potential)
                    if(length(potential)>0){
                        if(length(potential)==1)partner=potential
                        else partner=sample(potential,1)
                        population[[ind]]$partner=partner
                        population[[partner]]$partner=ind

                        if( population[[ind]]$sex == "F"){
                            fc=population[[ind]]$community
                            mc=population[[partner]]$community
                            stopifnot(population[[partner]]$sex=="M") #temp test
                        }
                        else{
                            mc=population[[ind]]$community
                            fc=population[[partner]]$community
                            stopifnot(population[[partner]]$sex=="F")

                        }
                        ## choice of new community  // ugly ; could do a function(rho,fc,mc) return(lc,jc) & ifelse
                        if(runif(1)<rho){
                            lc=fc #leaving the father's community
                            jc=mc #joining the mother's community
                        }
                        else{
                            jc=fc #joining the father's community
                            lc=mc #leaving the mother's community
                        }

                        initcomus$size[jc]=initcomus$size[jc]+1
                        initcomus$size[lc]=initcomus$size[lc]-1
                        population[[ind]]$community=population[[partner]]$community=jc

                        if(logging=="pairing")print(paste("marriage",ind,partner,"moving all to",jc," and leaving",lc, ",new:" ,population[[partner]]$community,population[[ind]]$community))
                    }
                }
            }

            #Handle deaths
            if(runif(1)<d){
                if(partner>0) population[[partner]]$partner=-1
                initcomus$size[population[[ind]]$community]=initcomus$size[population[[ind]]$community]-1
                if(logging=="demo") print(paste("dead ind",ind,"in",population[[ind]]$community))
                population[[ind]]=NULL
            }
        }

        ##pm transmission

    }

    ## Fission
    overloaded=initcomus$size>Th
    if(sum(overloaded)>0){
        listcomu=sapply(population,"[[","community")
        for(split in which(overloaded)){
            print(paste("community",split,"with size",initcomus$size[split]))
            tosplit= (listcomu == split)
            initcomus
            moving=sample(names(population)[tosplit],initcomus$size[split]/2)
        }

    }

    #quick summary of population at time 
    popsum[[time]]=sapply(names(population[[1]]),function(n)table(sapply(population,"[[",n)))

    ##
    plot(initcomus$coordinates,pch=21,bg=apply(initcomus$adaptivetraits,1,mean)+1,cex=log(initcomus$size))
    popsize=c(popsize,length(population))

    stopifnot(any(initcomus$size==table(sapply(population,"[[","community"))))
}

plot(popsize)

sumcomu=sapply(popsum,"[[","community")
plot(1,1,ylim=range(0,sumcomu),xlim=c(0,ncol(sumcomu)),type="n")
lapply(1:K,function(k) lines(sumcomu[k,],col=mean(initcomus$adaptivetraits[k,])+1,lwd=2) )

>>>>

devtools::load_all(".")
N=200
Th=400 #fission threshold
ki <- 1
km <- 1
K  <-  ki+km
m=1 #proba marriage
b=.2 #HG birth rate
r=.05 #farmer scale param
rho=.5 #marriage rule
d=0.001
maturity=0
endrepro=20
tstep=20*10

a=initAdaptiveTraits(ki=ki,km=km)
z=4 #number of neutral traits
neutraltraitsParam=initNeutralTraits(z = z)
neutraltraitsParam$s=c(0,0,1,1)

@@

pos=random2Dgrid(K=K,Gx=100)
initcomus=initialiseCommunities(traits=a,coordinates=pos)
size=table(sapply(population,"[[","community"))
initcomus$size=size
plot(initcomus$coordinates,pch=21,bg=apply(initcomus$adaptivetraits,1,mean)+1,cex=log(initcomus$size))

tentries=replicate(10,modelOOstyle( N=N, Th=Th, K =K, m=m, b=b, r=r, rho=rho, d=d, maturity=maturity, endrepro=endrepro, logging="time", population=population, initcomus=initcomus, tstep=tstep, neutraltraitsParam=neutraltraitsParam))

plot(1:200,ylim=c(0,600),type="n")
popsums=apply(tentries,2,function(res){
                   popsum=res$popsum
                   sumcomu=sapply(popsum,"[[","community")
                   lapply(1:K,function(k) lines(sumcomu[k,],col=mean(initcomus$adaptivetraits[k,])+1,lwd=2) )
})

lines(1:length(fa)*(endrepro-maturity),fa,col="red",lty=2,lwd=2)
lines(1:length(hg)*(endrepro-maturity),hg,col="black",lty=3,lwd=2)

moveA <- function(listmove,pop,newcom){
            for( m in listmove){
                pop[[m]]$community = newcom
            }
}

moveB <- function(listmove,pop,newcom){
    pop[moving]=lapply(pop[moving],function(m,nc){m$commnunity=nc;m},nc=newcom)
}

library(microbenchmark)
microbenchmark(moveA(moving,population,2),moveB(moving,population,2))

#keep horizontal: +/- 10 years  oblique

#when list (remove and fission)
plot(1,1,ylim=range(0,sumcomu),xlim=c(0,max(lengths(sumcomu))),type="n")
lapply(1:K,function(k) lines(sumcomu[[k]]),col=mean(initcomus$adaptivetraits[k,])+1,lwd=2) )
